# pre-class


Make sure you commit this often with meaningfull messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 


```{r}
library(plyr)
library(dplyr)
library(gapminder)
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
unique_country<-gapminder %>%
  group_by(continent) %>%
  distinct(country) %>%
  arrange(continent)
colSums(table(unique_country))
```
#There are 52 unique countries in Africa, 25 in the Americas, 33 in Asia, 30 in Europe, and 2 in Oceania. 

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
gapminder %>%
  filter(continent == "Europe", year=="1997") %>%
  arrange(gdpPercap)
```
#Albania had the lowest GDP per capita in 1997. 

3. According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
gapminder %>%
  group_by(continent) %>%
  filter(year >= "1980", year <= "1989") %>%
  summarize(avg_le = mean(lifeExp))
```
#In the 1980s, the average life expenctancy was 52.49 in Africa, 67.16 in the Americas, 63.73 in Asia, 73.22 in Europe, and 74.81 in Oceania. 

4. What 5 countries have the highest total GDP over all years combined?

```{r}
gapminder %>%
  group_by(country) %>%
  summarise(sum_GDP = sum(gdpPercap)) %>%
  arrange(desc(sum_GDP))
```
#The five countries with the highest total GDP over all years combined are Kuwait, Switzerland, Norway, United States, and Canada. 

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

```{r}
gapminder %>%
  group_by(country) %>%
  filter(lifeExp >= "80") %>%
  select(country, lifeExp, year)
```
#The countries with life expectancies over 80 are the following: Australia (2002,2007), Canada (2007), France (2007), Hong Kong, China (1997,2002,2007), Iceland (2002,2007), Israel (2007), Italy (2002,2007), Japan (1997,2002.2007), New Zealand (2007), Norway (2007), Spain (2007), Sweden (2002,2007), and Switzerland (2002,2007). 

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?

```{r}
xx <- data.frame(group = gapminder$country, a = gapminder$lifeExp , b = gapminder$gdpPercap)
head(xx)
require(plyr)
func <- function(xx)
return(data.frame(COR = cor(xx$a, xx$b)))
cor<-ddply(xx, .(group), func)
arrange(cor, desc(COR))
```
#The 10 countries with the strongest correlation between life expectancy and per capita gdp are France, Austria, Belgium, Norway, Oman, United Kingdom, Italy, Israel, Denmark, and Australia.  

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).

```{r}
gapminder %>%
  filter(continent != "Asia") %>%
  summarise(avg_pop = mean(pop)) %>%
  ungroup() %>%
  arrange(desc(avg_pop))
```

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 

```{r}
gapminder %>%
  group_by(country,pop)%>%
  distinct(country) %>%
  mutate(sd = sd(pop)) %>%
  arrange(sd)
```

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?

```{r}
gm1992<- gapminder %>%
  filter(year == "1992")
typeof(gm1992)
```
#This is a list that only includes observations for the year 1992. 

10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).

```{r}

```
