Katie Gately 

# pre-class

Make sure you commit this often with meaningful messages. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix  #Extract or replace substrings in a character vector
}                       #nchar returns a vector of prefix
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)  #If x is less than 1, it returns NULL
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))  #Replicates values in y, output vector is length x 
}
```
Renaming ideas:
f1: vector_prefix
f2: abs_value
f3: replace_x_y

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

```{r}
library(MASS)
?rnorm() #random variation of normal distribution: mean=0, sd=1
?mvrnorm() #1+ samples from specified normal distribution
```
#Rnorm and mvrnorm both generate normal distributions. However, mvrnorm allows you to specify the variance through the Sigma argument, whereas rnorm takes a default standard deviation of 1. If you specify a different sd in rnorm, the two functions may be more consistent. 

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}
lapply(mtcars,function(x) (sd(x)/mean(x)))
```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.

```{r}
df<-data.frame("age"=3:5,"date"=3:5)
df
vapply(df,function(y) {sd(y)}, numeric(1),USE.NAMES = TRUE)
```

    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
    
```{r}
mixed_df<-data.frame("weight"=2.4:6, "height" = "cm")
mixed_df
vapply(mixed_df[vapply(mixed_df, is.numeric, logical(1))], sd, numeric(1))
```

